Fitting correlated interferometric data

### 2021-01-26 article.tex
+ p2vm

@@

\section{Single value fit}

@@

\section{Model fitting}
We now consider a set of measurements corresponding to the linear model 
\begin{align}
    \vmod  &= \msens\vparam,\\
\intertext{where $\vparam$ are the unknown parameters and $\msens$ is the known sensitivity matrix. Typically, $\sens_{ik}=f_k(u_i, v_i)$ for a linear model and $\sens_{ik} = \partial f/\partial p_k(u_i, v_i)$ for a non-linear model that we choose to approximate by a linear one close to a solution. $(u, v)$ is the reduced baseline.  The true values $\vdatamean$ are are impacted by errors so that the data are} 
    \vdata &= \vdatamean + \verror\\
\intertext{with the error term $\verror$ again expressed as the sum of a measurement error and a normalisation one:}
    \verror &=  \vabserror + \vrelerror \hadam \vdatamean.
\end{align}
The measurement errors $\vec\abserror$ and normalisation errors $\vec\relerror$ following multivariate distributions of mean zero and covariance matrices $\vabscov$ and $\vrelcov$ respectively. Given the covariance matrix $\vcov$ of this model, the least squares estimate is
\begin{align}
   \vparam &= (\tr\msens \vcov^{-1} \msens)^{-1} (\tr\msens \vcov^{-1} \vdata)
\end{align}
We investigate four ways to determine the covariance matrix
\begin{enumerate}
    \item Ignoring the correlations in the normalisation using $\vcov_0 = \vabscov + (\vdata\outer\vdata)\hadam(\vrelcov\hadam\vec I)$. We note $\vdataflat = \msens(\tr\msens\vcov_0^{-1}\msens)^{-1}\tr\msens\vcov_0^{-1}\vdata$ the resulting model of the data. 
    \item Using the na\"\i{}ve estimate $\ppp\vcov = \vabscov + (\vdata\outer\vdata)\hadam\vrelcov$ which is known to lead to Peelle's pertinent puzzle in the trivial case of a constant model. 
    \item Using the data model of the fit without the normalisation error: $\vcov_1 = \vabscov + (\vdataflat\outer\vdataflat)\hadam\vrelcov$.  The resulting least squares model is $\vdatarec{1} = \msens(\tr\msens\vcov_1^{-1}\msens)^{-1}\tr\msens\vcov_1^{-1}\vdata$.
    \item Recursively fitting the data by updating the data model in the covariance matrix. We derive $\vdatarec{k} = \msens(\tr\msens\vcov_k^{-1}\msens)^{-1}\tr\msens\vcov_k^{-1}\vdata$ using $\vcov_{k} = \vabscov + (\vdatarec{k-1}\outer\vdatarec{k-1})\hadam\vrelcov$, starting with the estimate $\vdatarec{1}$ ($k = 2$). 
\end{enumerate}

@@

>>>>

\section{Normalisation by a transfer function}

\subsection{Reframing Peelle's Pertinent Puzzle}

@@

\subsection{Single value fit}

@@

\subsection{Model fitting}
We now consider a set of measurements corresponding to the linear model 
\begin{align}
    \vmod  &= \msens\vparam,\\
\intertext{where $\vparam$ are the unknown parameters and $\msens$ is the known sensitivity matrix. Typically, $\sens_{ik}=f_k(u_i, v_i)$ for a linear model and $\sens_{ik} = \partial f/\partial p_k(u_i, v_i)$ for a non-linear model that we choose to approximate by a linear one close to a solution. $(u, v)$ is the reduced baseline.  The true values $\vdatamean$ are are impacted by errors so that the data are} 
    \vdata &= \vdatamean + \verror\\
\intertext{with the error term $\verror$ again expressed as the sum of a measurement error and a normalisation one:}
    \verror &=  \vabserror + \vrelerror \hadam \vdatamean.
\end{align}
The measurement errors $\vec\abserror$ and normalisation errors $\vec\relerror$ following multivariate distributions of mean zero and covariance matrices $\vabscov$ and $\vrelcov$ respectively. Given the covariance matrix $\vcov$ of this model, the least squares estimate is
\begin{align}
   \vparam &= (\tr\msens \vcov^{-1} \msens)^{-1} (\tr\msens \vcov^{-1} \vdata)
\end{align}
We investigate four ways to determine the covariance matrix
\begin{enumerate}
    \item Ignoring the correlations in the normalisation using $\vcov_0 = \vabscov + (\vdata\outer\vdata)\hadam(\vrelcov\hadam\vec I)$. We note $\vdataflat = \msens(\tr\msens\vcov_0^{-1}\msens)^{-1}\tr\msens\vcov_0^{-1}\vdata$ the resulting model of the data. 
    \item Using the na\"\i{}ve estimate $\ppp\vcov = \vabscov + (\vdata\outer\vdata)\hadam\vrelcov$ which is known to lead to Peelle's pertinent puzzle in the trivial case of a constant model. 
    \item Using the data model of the fit without the normalisation error: $\vcov_1 = \vabscov + (\vdataflat\outer\vdataflat)\hadam\vrelcov$.  The resulting least squares model is $\vdatarec{1} = \msens(\tr\msens\vcov_1^{-1}\msens)^{-1}\tr\msens\vcov_1^{-1}\vdata$.
    \item Recursively fitting the data by updating the data model in the covariance matrix. We derive $\vdatarec{k} = \msens(\tr\msens\vcov_k^{-1}\msens)^{-1}\tr\msens\vcov_k^{-1}\vdata$ using $\vcov_{k} = \vabscov + (\vdatarec{k-1}\outer\vdatarec{k-1})\hadam\vrelcov$, starting with the estimate $\vdatarec{1}$ ($k = 2$). 
\end{enumerate}

@@

\section{Pixel-to-visibility matrix}
\def\vcovmat#1{\vec\Sigma^{\vec{#1}}}
\def\vcorrmat#1{\vec\rho^{\vec{#1}}}
\def\covmat#1{\Sigma^{\vec{#1}}}
\def\corrmat#1{\rho^{\vec{#1}}}
I call $I_p$ the luminous intensity on detector part $p$ ($1 \cdots P$).  The scalar interferometric observable $V_o$ for observable $o$ ($1 \cdots O$) is deduced using the pixel-to-visibility matrix $M_{op}$ by
\begin{align}
    \vec{V} &= \vec{M}\vec{I},
\intertext{where $\vec{M}$ and $\vec{I}$ can be written as}
    \vec{I} &= \mean{\vec{I}}(1 + \vec\error^{\vec I}),\\
    \vec{M} &= \mean{\vec{M}}(1 + \vec\error^{\vec M})
\end{align}

The covariance matrix $\covmat{V}$ can be obtained from corralation matrices 
\begin{align}
    \vcorrmat{I} &= <\vec\error^{\vec M} \outer \vec\error^{\vec I}>\\
    \vcorrmat{M} &= <\vec\error^{\vec M} \outer \vec\error^{\vec M}>
\end{align}
as
\begin{equation}
   \begin{split}
   \covmat{V}_{oo'} = \sum_{1\le p,p'\le P}&
        \bigg[\Big( 
             \corrmat{I}_{pp'}
           + \corrmat{M}_{opop'}
           - \corrmat{I}_{pp'}\corrmat{M}_{opop'}
        \Big) \times\\
        &\qquad \mean{I}_p \mean{I}_{p'} \mean{M}_{op} \mean{M}_{o'p'}\bigg], 
    \end{split}
\end{equation}
assuming that $M$ and $I$ are not correlated.

Once again the true value $\mean{\vec{I}}$ is unknown and replacing it by the
observed value $\vec{I}$ may lead to bias. Unlike in the previous section,
$\vec{I}$ is not directly produced by the modelling, so one would need
to obtain it from $\vec{V}$.  However, the pixel-to-visibility matrix is
not necessarily invertible.
